{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-03-01 Paper_ttbarH_Draft.tex
Changes discussed in skype meeting

%
Before turning to a discussion of $\mathrm{CP}$-odd observables,
it is useful to consider a few theoretical aspects of
the process $\ppprocess$.  In this section we derive a ``factorized''
expression for the gluon fusion contribution to this process.
We then use this expression to isolate various mathematical quantities
that will be useful as we construct $\mathrm{CP}$-odd observables.

\subsection{Factorized expression for the scattering cross section}

At the LHC $t\bar{t}H$ production proceeds via $q\bar{q}$
annihilation and $gg$ fusion processes. The relevant leading-order Feynman
diagrams are displayed in Fig.~\ref{fig1}, where the first two
rows show the $q\bar{q}$ and $gg$ $s$-channel diagrams, and
the last one depicts the $gg$ $t$-channel diagrams.  Three more
$gg$-initiated diagrams are obtained by exchanging the gluon
lines in the third row.
We describe the $tH$ coupling with the effective Lagrangian
\label{eq1}
\mathcal{L}_{t\bar{t}H}=-\frac{m_t}{v}(\kp \tbar t+i\kpt \tbar
\gamma_5 t)H,
where $v=246~\mathrm{GeV}$ is the SM Higgs vacuum
expectation value, and the coefficients $\kp$ and $\kpt$ parameterize 
the scalar and pseudoscalar interaction, respectively. The
SM case is obtained for $\kp=1$ and $\kpt=0$, while the values $\kp=0$
and $\kpt\neq 0$ parameterize a $\mathrm{CP}$-odd Higgs boson. \par
%%%%%%%%%%%%%%%%%%%% Figura 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/home/nico/jaxodraw-2.1-0/
\centering
%\hspace*{-0.4cm}
\subfloat{\includegraphics[scale=0.45]{qqs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1a}}
\subfloat{\includegraphics[scale=0.45]{qqs2_II.pdf}}
%\label{fig1b}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1b}}
\subfloat{\includegraphics[scale=0.45]{ggs2_II.pdf}}
%\label{fig1c}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggt1_II.pdf}}
%\label{fig1b}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt2_II.pdf}}
%\label{fig1c}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt3_II.pdf}}
%\label{fig1b}}
\vspace*{0.02\textwidth}
\caption{Tree-level Feynman diagrams contributing to $t\tbar H$ production
  at the LHC. Three more diagrams are obtained by exchanging the gluon
  lines in the $t$-channel diagrams.}
\label{fig1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this section we focus on the gluon fusion contributions
to $t\tbar H$ production, since these dominate over the
the quark-antiquark annihilation contributions.  Furthermore,
we consider the case in which the top and antitop both decay
semileptonically.  As we shall show below, assuming the
narrow width approximation for the top and antitop quarks, the unpolarized
differential cross section for $gg\to t(\to
b{\ell}^+\nu_{\ell})\,\tbar(\to \bbar {\ell}^- \nubar)\,H$ may be
written in the following ``factorized'' form,\footnote{The reader
  is referred to the discussion following Eq.~(\ref{eq17}) for some qualifying
  remarks regarding the ``factorization'' of this expression.}
\label{eq2}
%dd\sigma(gg\to (bl^+\nu_l)(\bbar l^- \nubar) H)=\sum_{\substack{bl^+\nu_l \\ \tiny{\mathrm{spins}}}}\,\sum_{\substack{\bbar l^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,d\Gamma(t(n_t)\to bl^+\nu_l)\,d\Gamma(\tbar (n_{\tbar})\to \bbar l^-\nubar)
d\sigma =\sum_{\substack{b{\ell}^+\nu_l \\ \tiny{\mathrm{spins}}}}\,
   \sum_{\substack{\bbar {\ell}^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,
   d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,
   d\Gamma(t\to b{\ell}^+\nu_{\ell})\,
   d\Gamma(\tbar \to \bbar {\ell}^-\nubar),
where $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$ is the differential
cross section for the production of a top and antitop quark,
with spin vectors $n_t$ and $n_{\bar{t}}$, respectively, along with a Higgs
boson.  Also, $d\Gamma(t\to b{\ell}^+\nu_{\ell})$ and
$d\Gamma(\tbar \to \bbar {\ell}^-\nubar)$ are the
partial differential decay widths for an unpolarized top and
anti-top quark.  The four-vectors
$n_t$ and $n_{\tbar}$ are not arbitrary, but are
given by particular combinations of the
momenta of the $t,\tbar, \ell^+$ and $\ell^-$~\cite{Arens},
\bea
\label{eq3}
n_t&=&-\frac{p_t}{m_t}+\frac{m_t}{(p_t\cdot p_{{\ell}^+})}p_{{\ell}^+}\\
\label{eq4}
n_{\tbar}&=&\,\frac{p_{\tbar}}{m_t}-\frac{m_t}{(p_{\tbar}\cdot p_{{\ell}^-})}p_{{\ell}^-}.
\eea
Expressions similar to Eq.~(\ref{eq2}) have been derived previously for the
production of short-lived particles in $e^-e^+$ colliders
\cite{kawasaki} and for $t\tbar$ production both in
$e^-e^+$ colliders \cite{Arens} and $pp$ colliders
\cite{ale1,*ale2,*ale3,*ale4}. \par

@@

To derive the above expressions, we begin by considering the
schematic representation for the process $g_ag_b\to t(\to
b_i{\ell}^+\nu_{\ell})\,\tbar(\to \bbar_j {\ell}^- \nubar)\,H$
that is sketched in Fig.~\ref{fig2}.
Here $a$ and $b$ denote the initial-state gluons and
$i$ and $j$ refer to the colours of the top and antitop quarks.
%\par
%
The amplitude for this process may be written in the following
compact form
%
\beq
\label{eq6}
\mathcal{M}^{ab,ij}=\bar{\psi}_t\,\mathcal{A}^{ab,ij}\,\psi_{\tbar}\,,
\eeq
%
where the spinors $\bar{\psi}_t$ and $\psi_{\tbar}$
contain all of the information regarding the decay of the
virtual top and anti-top, respectively,
and where the quantity $A^{ab,ij}$ is given by
%
\beq
\label{eq5}
\mathcal{A}^{ab,ij}\equiv A^{ab,ij}_{\mu\nu}(\epsilon_{\lambda_a})^{\mu}(\epsilon_{\lambda_b})^{\nu}=\sum_{k=1}^8 \mathcal{A}^{ab,ij}_k =\kp \sum_{k=1}^8 \mathcal{S}^{ab,ij}_k + i\kpt \sum_{k=1}^8 \mathcal{P}^{ab,ij}_k.
\eeq 
%
The sum over $k$ in the above expression
corresponds to the eight gluon-initiated
diagrams indicated in Fig.~\ref{fig1}; also,
$\epsilon_{\lambda_a}$ and $\epsilon_{\lambda_b}$ are the
polarization vectors corresponding to $g_a$ and $g_b$, respectively.
In the last equality in Eq.~(\ref{eq5})
we have explicitly separated the amplitude into
two sums, with one sum corresponding to the scalar contributions
and the other to the pseudoscalar ones.
Taking all of the final-state particles to be massless, we can use the
spinor techniques
developed in \cite{Kleiss} to write $\bar{\psi}_t$ and
$\psi_{\tbar}$ as follows\footnote{These spinor techniques can also be used
  for massive final-state particles. Given the
  energy scale involved in the process in question, however, the assumption
  of massless final-state particles is sensible and greatly
  simplifies the derivation of Eq.~(\ref{eq2}).}
%
\bea
\label{eq7}
\bar{\psi}_t &=& -g^2\, \mathbb{P}_t(t)\,\mathbb{P}_W(t-b)\,
   \langle b-|\nu_{\ell}+\rangle \langle {\ell}^+\!+|(\tslash+m_t)\\
\label{eq8}
\psi_{\tbar} &=& g^2\, \mathbb{P}_t(\tbar)\,\mathbb{P}_W(\tbar-\bbar)\,
   \langle\nubar+|\bbar-\rangle (\tbslash-m_t)|{\ell}^-+\rangle,
\eea
%
where $|i+(-)\rangle \equiv (1/2)(1\pm \gamma^5)\,\psi_i$
represents a right-handed (left-handed) chiral spinor for final-state
particle $i$ and $\langle i+(-)|$ represents the corresponding adjoint spinor.
{\bf (KK: Do we mean the adjoint spinor, related by $C \psi^T$,
  or whatever?, N: It is $\psi^{\dagger}\gamma^0$. I'm using the name given in Eq(2.28) of Bjorken and Drell's book.)}
Also, $\mathbb{P}_t(q)=(q^2-m^2_t+im_t\Gamma_t)^{-1}$ and
$\mathbb{P}_W(q)=(q^2-m^2_W+im_W\Gamma_W)^{-1}$, and
we have denoted the momenta of the various particles by the symbols
that refer to the names of those particles~\cite{Mangano}.

Using the expressions defined above for $\bar{\psi}_t$ and
$\psi_{\tbar}$, we can write the amplitude $\mathcal{M}^{ab,ij}$
in a form that is (in a sense) factorized.
As a first step, we insert
Eqs.~(\ref{eq7}) and (\ref{eq8}) into Eq.~(\ref{eq6}), yielding
%
\beq
\label{eq9}
\mathcal{M}^{ab,ij}=\!-g^4\,\mathbb{P}_t(t)\,\mathbb{P}_t(\tbar)\,\mathbb{P}_W(t-b)\,\mathbb{P}_W(\tbar-\bbar)\,\langle b-|\nu_{\ell}+\rangle \langle\nubar+|\bbar-\rangle \sqrt{\strut2(t\cdot {\ell}^+)}\sqrt{\strut2(\tbar\cdot {\ell}^-)}\,[\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}],
%\frac{1}{(t^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{(\tbar^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{((t-b)^2-M^2_W)}\frac{1}{((\tbar-\bbar)^2-M^2_W)}\,\langle b-|\nu+\rangle\,\langle\nubar+|\bbar-\rangle
\eeq
%
where the spinors $\phi_{t}$ and $\phi_{\tbar}$ are defined as
%
\beq
\label{eq10}
\phi_t=\frac{(\tslash +m_t)}{\sqrt{\strut2(t\cdot {\ell}^+)}}|{\ell}^++\rangle
\eeq
%
\beq
\label{eq11}
\phi_{\tbar}=\frac{(\tbslash -m_t)}{\sqrt{\strut2(\tbar\cdot {\ell}^-)}}|{\ell}^-+\rangle \,.
\eeq
%
Note that in writing down the above expressions we have adopted the narrow-width
approximation for the top and antitop quarks and for the $W^\pm$ gauge
bosons.\footnote{Since Eq.~(\ref{eq9}) contains the top quark propagator
  term
  $\mathbb{P}_t(t)$, for example, $|\mathcal{M}^{ab,ij}|^2$ contains
  the factor $((t^2-m^2_t)^2+m^2_t\Gamma^2_t)^{-1}$,
  which is replaced by $(\pi/m_t\Gamma_t)\delta(t^2-m^2_t)$
  in the narrow-width approximation.  Thus we take the four-vector
  $t$ appearing in the definition of $\phi_t$ to be
  on shell, satisfying $t^2 = m_t^2$.} {\bf (KK: Do I understand correctly
  that $t$ is an on-shell four-vector in $t\cdot \ell^+$, but not
  in $\mathbb{P}_t(t)$?  This is similar to saying $f(x)\delta(x-x_0) =
  f(x_0)\delta(x-x_0)$\ldots?)}
Working out the projection operators $\phi_t\,\bar{\phi}_t$
and $\phi_{\tbar}\,\bar{\phi}_{\tbar}$, we have
%
\beq
\label{eq12}
\phi_t\,\bar{\phi}_t=\frac{1}{2}(1+\nslash_{\!t}\gamma^5)(\tslash +m_t)
\eeq
%
and
%
\beq
\label{eq13}
\phi_{\tbar}\,\bar{\phi}_{\tbar}=\frac{1}{2}(1+\nslash_{\!\tbar}\gamma^5)(\tbslash -m_t),
\eeq
%
with $n_t$ and $n_{\tbar}$ being the four-vectors defined in
Eqs.~(\ref{eq3}) and (\ref{eq4}).  Thus, $\phi_t$ and $\phi_{\tbar}$
may be regarded as
describing a top quark with spin vector $n_t$ and an antitop quark
with spin vector $n_{\tbar}$, respectively.

As a final step toward factorizing the amplitude $\mathcal{M}^{ab,ij}$,
we note that the amplitude for a top quark with spin vector
$n_t$ to decay into $b{\ell}^+\nu_{\ell}$ is given by
%
\beq
\label{eq14}
\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})=ig^2\mathbb{P}_W(t-b)\langle b-|\nu_{\ell}+\rangle\sqrt{\strut2(t\cdot {\ell}^+)} \,,
\eeq
%
and likewise,
%
\beq
\label{eq15}
\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})=ig^2\mathbb{P}_W(\tbar-\bar{b})\langle \bar{\nu}_{\ell}+|\bar{b}-\rangle\sqrt{\strut2(\tbar\cdot {\ell}^-)}.
\eeq
%
Furthermore, the term inside the square brackets in
Eq.~(\ref{eq9}) is the amplitude for producing a top quark
with spin vector $n_t$, along with an anti-top with spin vector
$n_{\tbar}$ and a Higgs boson,
%
\beq
\label{eq16}
\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)=\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}.
\eeq
%
Combining Eqs.~(\ref{eq14})-(\ref{eq16}), we can write Eq.~(\ref{eq9})
in a form that appears to be factorized,
%
\beq
\label{eq17}
\mathcal{M}^{ab,ij}=\mathbb{P}_t(t)\mathbb{P}_t(\tbar)\,\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})\,\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})\,\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H) \,.
\eeq
%
It is important
to note that, even though the above expression has the appearance of being
factorized into production and decay parts, this apparent
factorization is a bit misleading.  In particular, the amplitude for
$t\tbar H$ production contains the top and antitop quark
spin four-vectors $n_t$ and $n_{\tbar}$, which
depend on final-state kinematical quantities [see
Eqs.~(\ref{eq3}) and (\ref{eq4})].
With this qualification in mind, we may now use the
amplitude in Eq.~(\ref{eq17})
to determine the corresponding scattering cross section.
After some manipulation of the phase space variables
to take advantage of the presence of the
propagator terms, $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
we arrive at the expression in Eq.~(\ref{eq2}).  This expression
also has the appearance of being factorized, but
qualifying remarks, similar to those above, apply.
{\bf (KK: I'm still a bit confused as to why $n_t$ appears
  in the amplitude expression for the decay, but not
  in the expression for the cross section.  That is, the top
  spins are summed over in the width that appears in the
  factorized expression for the cross section.  That part makes
  sense to me, but I'm not so clear on how to get there from
  Eq.~(17).  Probably need to look more closely at the notes!)}

@@

\subsection{Triple products}
{\bf (KK: maybe we should change this title for the subsection\ldots.)}

@@

$\kp$ and $\kpt$, they {\bf are} expected to vanish in these cases. However, we

>>>>

%
Before turning to a discussion of $\mathrm{CP}$-odd observables,
it is useful to consider a few theoretical aspects of
the process $\ppprocess$.  In this section we derive a ``factorized''
expression for the gluon fusion contribution to this process.
We then use this expression to isolate various mathematical quantities
that will be useful as we construct $\mathrm{CP}$-odd observables.\par

At the LHC $t\bar{t}H$ production proceeds via $q\bar{q}$
annihilation and $gg$ fusion processes. The relevant leading-order Feynman
diagrams are displayed in Fig.~\ref{fig1}, where the first two
rows show the $q\bar{q}$ and $gg$ $s$-channel diagrams, and
the last one depicts the $gg$ $t$-channel diagrams.  Three more
$gg$-initiated diagrams are obtained by exchanging the gluon
lines in the third row.
We describe the $tH$ coupling with the effective Lagrangian
%
\beq
\label{eq1}
\mathcal{L}_{t\bar{t}H}=-\frac{m_t}{v}(\kp \tbar t+i\kpt \tbar
\gamma_5 t)H,
\eeq
%
where $v=246~\mathrm{GeV}$ is the SM Higgs vacuum
expectation value, and the coefficients $\kp$ and $\kpt$ parameterize 
the scalar and pseudoscalar interaction, respectively. The
SM case is obtained for $\kp=1$ and $\kpt=0$, while the values $\kp=0$
and $\kpt\neq 0$ parameterize a $\mathrm{CP}$-odd Higgs boson. \par
%%%%%%%%%%%%%%%%%%%% Figura 1%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/home/nico/jaxodraw-2.1-0/
\begin{center}
\begin{figure}[H]
\centering
%\hspace*{-0.4cm}
\subfloat{\includegraphics[scale=0.45]{qqs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1a}}
\subfloat{\includegraphics[scale=0.45]{qqs2_II.pdf}}
%\label{fig1b}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggs1_II.pdf}}
\hspace*{0.05\textwidth}
%\label{fig1b}}
\subfloat{\includegraphics[scale=0.45]{ggs2_II.pdf}}
%\label{fig1c}}
\\[0.032\textwidth]
\subfloat{\includegraphics[scale=0.45]{ggt1_II.pdf}}
%\label{fig1b}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt2_II.pdf}}
%\label{fig1c}}
\hspace*{0.025\textwidth}
\subfloat{\includegraphics[scale=0.45]{ggt3_II.pdf}}
%\label{fig1b}}
\vspace*{0.02\textwidth}
\caption{Tree-level Feynman diagrams contributing to $t\tbar H$ production
  at the LHC. Three more diagrams are obtained by exchanging the gluon
  lines in the $t$-channel diagrams.}
\label{fig1}
\end{figure}
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Factorized expression for the scattering cross section}

In this subsection we focus on the gluon fusion contributions
to $t\tbar H$ production, since these dominate over the
the quark-antiquark annihilation contributions.  Furthermore,
we consider the case in which the top and antitop both decay
semileptonically.  As we shall show below, assuming the
narrow width approximation for the top and antitop quarks, the unpolarized
differential cross section for $gg\to t(\to
b{\ell}^+\nu_{\ell})\,\tbar(\to \bbar {\ell}^- \nubar)\,H$ may be
written in the following ``factorized'' form,\footnote{The reader
  is referred to the discussion following Eq.~(\ref{eq17}) for some qualifying
  remarks regarding the ``factorization'' of this expression.}
%
\beq
\label{eq2}
%dd\sigma(gg\to (bl^+\nu_l)(\bbar l^- \nubar) H)=\sum_{\substack{bl^+\nu_l \\ \tiny{\mathrm{spins}}}}\,\sum_{\substack{\bbar l^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,d\Gamma(t(n_t)\to bl^+\nu_l)\,d\Gamma(\tbar (n_{\tbar})\to \bbar l^-\nubar)
d\sigma =\sum_{\substack{b{\ell}^+\nu_l \\ \tiny{\mathrm{spins}}}}\,
   \sum_{\substack{\bbar {\ell}^-\nubar \\ \mathrm{\tiny{spins}}}}\left(\frac{2}{\Gamma_t}\right)^2\,
   d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)\,
   d\Gamma(t\to b{\ell}^+\nu_{\ell})\,
   d\Gamma(\tbar \to \bbar {\ell}^-\nubar),
\eeq  
%
where $d\sigma(gg\to t(n_t)\tbar (n_{\tbar})H)$ is the differential
cross section for the production of a top and antitop quark,
with spin vectors $n_t$ and $n_{\bar{t}}$, respectively, along with a Higgs
boson.  Also, $d\Gamma(t\to b{\ell}^+\nu_{\ell})$ and
$d\Gamma(\tbar \to \bbar {\ell}^-\nubar)$ are the
partial differential decay widths for an unpolarized top and
anti-top quark.  The four-vectors
$n_t$ and $n_{\tbar}$ are not arbitrary, but are
given by particular combinations of the
momenta of the $t,\tbar, \ell^+$ and $\ell^-$~\cite{Arens},
%
\bea
\label{eq3}
n_t&=&-\frac{p_t}{m_t}+\frac{m_t}{(p_t\cdot p_{{\ell}^+})}p_{{\ell}^+}\\
\label{eq4}
n_{\tbar}&=&\,\frac{p_{\tbar}}{m_t}-\frac{m_t}{(p_{\tbar}\cdot p_{{\ell}^-})}p_{{\ell}^-}.
\eea
%
Expressions similar to Eq.~(\ref{eq2}) have been derived previously for the
production of short-lived particles in $e^-e^+$ colliders
\cite{kawasaki} and for $t\tbar$ production both in
$e^-e^+$ colliders \cite{Arens} and $pp$ colliders
\cite{ale1,*ale2,*ale3,*ale4}. \par

@@

To derive the above expressions, we begin by considering the
schematic representation for the process $g_ag_b\to t(\to
b_i{\ell}^+\nu_{\ell})\,\tbar(\to \bbar_j {\ell}^- \nubar)\,H$
that is sketched in Fig.~\ref{fig2}.
Here $a$ and $b$ denote the initial-state gluons and
$i$ and $j$ refer to the colours of the top and antitop quarks.
%\par
%
The amplitude for this process may be written in the following
compact form
%
\beq
\label{eq6}
\mathcal{M}^{ab,ij}=\bar{\psi}_t\,\mathcal{A}^{ab,ij}\,\psi_{\tbar}\,,
\eeq
%
where the spinors $\bar{\psi}_t$ and $\psi_{\tbar}$
contain all of the information regarding the decay of the
virtual top and anti-top, respectively,
and where the quantity $A^{ab,ij}$ is given by
%
\beq
\label{eq5}
\mathcal{A}^{ab,ij}\equiv A^{ab,ij}_{\mu\nu}(\epsilon_{\lambda_a})^{\mu}(\epsilon_{\lambda_b})^{\nu}=\sum_{k=1}^8 \mathcal{A}^{ab,ij}_k =\kp \sum_{k=1}^8 \mathcal{S}^{ab,ij}_k + i\kpt \sum_{k=1}^8 \mathcal{P}^{ab,ij}_k.
\eeq 
%
The sum over $k$ in the above expression
corresponds to the eight gluon-initiated
diagrams indicated in Fig.~\ref{fig1}; also,
$\epsilon_{\lambda_a}$ and $\epsilon_{\lambda_b}$ are the
polarization vectors corresponding to $g_a$ and $g_b$, respectively.
In the last equality in Eq.~(\ref{eq5})
we have explicitly separated the amplitude into
two sums, with one sum corresponding to the scalar contributions
and the other to the pseudoscalar ones.
Taking all of the final-state particles to be massless, we can use the
spinor techniques
developed in \cite{Kleiss} to write $\bar{\psi}_t$ and
$\psi_{\tbar}$ as follows\footnote{These spinor techniques can also be used
  for massive final-state particles. Given the
  energy scale involved in the process in question, however, the assumption
  of massless final-state particles is sensible and greatly
  simplifies the derivation of Eq.~(\ref{eq2}).}
%
\bea
\label{eq7}
\bar{\psi}_t &=& -g^2\, \mathbb{P}_t(t)\,\mathbb{P}_W(t-b)\,
   \langle b-|\nu_{\ell}+\rangle \langle {\ell}^+\!+|(\tslash+m_t)\\
\label{eq8}
\psi_{\tbar} &=& g^2\, \mathbb{P}_t(\tbar)\,\mathbb{P}_W(\tbar-\bbar)\,
   \langle\nubar+|\bbar-\rangle (\tbslash-m_t)|{\ell}^-+\rangle,
\eea
%
where $|i+(-)\rangle \equiv (1/2)(1\pm \gamma^5)\,\psi_i$
represents a right-handed (left-handed) chiral spinor for final-state
particle $i$ and $\langle i+(-)|$ represents the corresponding adjoint spinor.
Also, $\mathbb{P}_t(q)=(q^2-m^2_t+im_t\Gamma_t)^{-1}$ and
$\mathbb{P}_W(q)=(q^2-m^2_W+im_W\Gamma_W)^{-1}$, and
we have denoted the momenta of the various particles by the symbols
that refer to the names of those particles~\cite{Mangano}.

Using the expressions defined above for $\bar{\psi}_t$ and
$\psi_{\tbar}$, we can write the amplitude $\mathcal{M}^{ab,ij}$
in a form that is (in a sense) factorized.
As a first step, we insert
Eqs.~(\ref{eq7}) and (\ref{eq8}) into Eq.~(\ref{eq6}), yielding
%
\beq
\label{eq9}
\mathcal{M}^{ab,ij}=\!-g^4\,\mathbb{P}_t(t)\,\mathbb{P}_t(\tbar)\,\mathbb{P}_W(t-b)\,\mathbb{P}_W(\tbar-\bbar)\,\langle b-|\nu_{\ell}+\rangle \langle\nubar+|\bbar-\rangle \sqrt{\strut2(t\cdot {\ell}^+)}\sqrt{\strut2(\tbar\cdot {\ell}^-)}\,[\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}],
%\frac{1}{(t^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{(\tbar^2-m^2_t+im_t\Gamma_t)}\,\frac{1}{((t-b)^2-M^2_W)}\frac{1}{((\tbar-\bbar)^2-M^2_W)}\,\langle b-|\nu+\rangle\,\langle\nubar+|\bbar-\rangle
\eeq
%
where the spinors $\phi_{t}$ and $\phi_{\tbar}$ are defined as
%
\beq
\label{eq10}
\phi_t=\frac{(\tslash +m_t)}{\sqrt{\strut2(t\cdot {\ell}^+)}}|{\ell}^++\rangle
\eeq
%
\beq
\label{eq11}
\phi_{\tbar}=\frac{(\tbslash -m_t)}{\sqrt{\strut2(\tbar\cdot {\ell}^-)}}|{\ell}^-+\rangle \,.
\eeq
%
Note that in writing down the above expressions we have adopted the narrow-width
approximation for the top and antitop quarks and for the $W^\pm$ gauge
bosons.\footnote{Since Eq.~(\ref{eq9}) contains the top quark propagator
  term
  $\mathbb{P}_t(t)$, for example, $|\mathcal{M}^{ab,ij}|^2$ contains
  the factor $((t^2-m^2_t)^2+m^2_t\Gamma^2_t)^{-1}$,
  which is replaced by $(\pi/m_t\Gamma_t)\delta(t^2-m^2_t)$
  in the narrow-width approximation.  Thus, except for the propagator terms $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
   we take the four-vector $t$ appearing in Eqs.~(\ref{eq9})-(\ref{eq11}) to be
  on shell, satisfying $t^2 = m_t^2$.} 
Working out the projection operators $\phi_t\,\bar{\phi}_t$
and $\phi_{\tbar}\,\bar{\phi}_{\tbar}$, we have
%
\beq
\label{eq12}
\phi_t\,\bar{\phi}_t=\frac{1}{2}(1+\nslash_{\!t}\gamma^5)(\tslash +m_t)
\eeq
%
and
%
\beq
\label{eq13}
\phi_{\tbar}\,\bar{\phi}_{\tbar}=\frac{1}{2}(1+\nslash_{\!\tbar}\gamma^5)(\tbslash -m_t),
\eeq
%
with $n_t$ and $n_{\tbar}$ being the four-vectors defined in
Eqs.~(\ref{eq3}) and (\ref{eq4}).  Thus, $\phi_t$ and $\phi_{\tbar}$
may be regarded as
describing a top quark with spin vector $n_t$ and an antitop quark
with spin vector $n_{\tbar}$, respectively.

As a final step toward factorizing the amplitude $\mathcal{M}^{ab,ij}$,
we note that the amplitude for a top quark with spin vector
$n_t$ to decay into $b{\ell}^+\nu_{\ell}$ is given by
%
\beq
\label{eq14}
\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})=ig^2\mathbb{P}_W(t-b)\langle b-|\nu_{\ell}+\rangle\sqrt{\strut2(t\cdot {\ell}^+)} \,,
\eeq
%
and likewise,
%
\beq
\label{eq15}
\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})=ig^2\mathbb{P}_W(\tbar-\bar{b})\langle \bar{\nu}_{\ell}+|\bar{b}-\rangle\sqrt{\strut2(\tbar\cdot {\ell}^-)}.
\eeq
%
Furthermore, the term inside the square brackets in
Eq.~(\ref{eq9}) is the amplitude for producing a top quark
with spin vector $n_t$, along with an anti-top with spin vector
$n_{\tbar}$ and a Higgs boson,
%
\beq
\label{eq16}
\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H)=\bar{\phi}_t \mathcal{A}^{ab,ij}\phi_{\tbar}.
\eeq
%
Combining Eqs.~(\ref{eq14})-(\ref{eq16}), we can write Eq.~(\ref{eq9})
in a form that appears to be factorized,
%
\beq
\label{eq17}
\mathcal{M}^{ab,ij}=\mathbb{P}_t(t)\mathbb{P}_t(\tbar)\,\mathcal{M}(t(n_t)\to b{\ell}^+\nu_{\ell})\,\mathcal{M}(\tbar(n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})\,\mathcal{M}(g_ag_b \to t^i(n_t)\tbar^j(n_{\tbar})H) \,.
\eeq
%
It is important
to note that, even though the above expression has the appearance of being
factorized into production and decay parts, this apparent
factorization is a bit misleading.  In particular, the amplitude for
$t\tbar H$ production contains the top and antitop quark
spin four-vectors $n_t$ and $n_{\tbar}$, which
depend on final-state kinematical quantities [see
Eqs.~(\ref{eq3}) and (\ref{eq4})].
With this qualification in mind, we may now use the
amplitude in Eq.~(\ref{eq17})
to determine the corresponding scattering cross section.
After some manipulation of the phase space variables
to take advantage of the presence of the
propagator terms, $\mathbb{P}_t(t)$ and $\mathbb{P}_t(\tbar)$,
we arrive at the expression in Eq.~(\ref{eq2}).\footnote{The reader may note that in the differential widths of $t\to b{\ell}^+\nu_{\ell}$ and $\tbar\to \bar{b}{\ell}^-\bar{\nu}_{\ell}$ appearing in 
Eq.~(\ref{eq2}), the spin states of the top an antitop have been averaged. Interestingly, under the assumption of massless final-state particles, the amplitudes $\mathcal{M}(t(-n_t)\to b{\ell}^+\nu_{\ell})$ and $\mathcal{M}(\tbar(-n_{\tbar})\to \bar{b}{\ell}^-\bar{\nu}_{\ell})$ vanish. } This expression
also has the appearance of being factorized, but
qualifying remarks, similar to those above, apply.

@@

\subsection{Origin of Triple product terms}

@@

$\kp$ and $\kpt$, they expected to vanish in these cases ({\bf ``..they are expected to.."? Maybe it's a typo? we are not sure}). However, we
